---
title: "experiments-Jirka"
output: html_document
date: "2025-07-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("functions.R")
```

- čísla otázek jsou uvedena v souboru Objektivita přírodních věd - I. část (MU).pdf v adresáři materials
- propojení názvů a popisu je v souboru názvy-a-popis.xlsx v adresáři data_and_R_src

## Petr Jedlička email z 25.2.2025

1) Reprodukovatelnost (Názor na krizi) 
- předpokládáme vliv:
a) Obor (dle angl. číselníku, který je v obou výzkumech stejný), 
b) dále Sociodemografické faktory
c) Akademická kariéra (Ot 41-48), případně i další otázky: č. 50 podíl základního výzkumu u respondenta, pří. Ot: 49 - podíl věd činnosti u respondenta, Ot 50 - podíl zákl/apl výzkumu, které by tam mohly spadat 
Metoda: MLR, multinomial log. regression?

After we call prepare-data.Rmd we further process variables:

Variables classification into following groups:
- one dependent variable "krize_rep"
- explanatory variables:
  - Likert
  - ordinal
  - Boolean
  - categorical
  - complicated (will be ignored)
  - other variables (will be ignored)

```{r}
my <- my_clean

dependent.var <- "krize_rep"
# If we want to make dependent variable ordinal
my[[dependent.var]] <- ordered(my[[dependent.var]],
                              levels = c("Ano, významná krize","Ano, nevýznamná krize","Nevím","Ne, žádná krize není"))
  barplot(table(my[,dependent.var]),cex.names = 0.8)

explanatory.vars <- c(
"spec_hlavni",
"gender",           
"vek",          
"kar_praxe",
"kar_titul",       
"kar_citace",      
"kar_nejlepsi",     
"kar_kategorie",
"pomer_veda",
"pomer_zakladni",
"obor_dotovany",
"obor_konkuren",
"obor_tlak",
"obor_rychly",
"obor_priliv",
"obor_publikac",
"obor_medializ",
"obor_verejnos",
"obor_komplexn",
"obor_mensi",
"cinnost_matem",
"cinnost_model",
"cinnost_teori",
"cinnost_exper",
"cinnost_kateg",
"cinnost_vyvoj",
"obor_ekzaktnost")

# Když máme "spec_hlavni" (10 kategorií) asi už nechceme "obor" (12 kategorií, navíc podivných) 
# Jak se liší Aplikovaná fyzika od Sekce aplikované fyziky? Třeba spojit do jedné.
# Podobně u ostatních - nakonec vede na 6 oborů (sekcí)
other.vars <- c("obor")

# Nyní specifikujeme, které vysvětlující veličiny explanatory.vars se mohou chápat pomocí Likertovy škály.
# Veličiny mají 5 odpovědí, ale Exaktnost má pouze 3 odpovědi.
# Jak zacházet s odpovědí "Nevím"? 
# Spíše než NA jako s neutrálním stavem?
# Je třeba stanovit uspořádání 
likert.vars <- c(
"obor_dotovany",
"obor_konkuren",
"obor_tlak",
"obor_rychly",
"obor_priliv",
"obor_publikac",
"obor_medializ",
"obor_verejnos",
"obor_komplexn",
"obor_mensi",
"obor_ekzaktnost"
)
for (i in 1:(length(likert.vars)-1)){
  my[[likert.vars[i]]] <- ordered(my[[likert.vars[i]]], 
                                  levels = c("Rozhodně ne", "Spíše ne", "Nevím", "Spíše ano", "Rozhodně ano"))
}
# the last Likert variable has different states
i <- length(likert.vars)
my[[likert.vars[i]]] <- ordered(my[[likert.vars[i]]], 
                                  levels = c("„Měkčí“ (méně exaktní)", "Nedokážu posoudit", "„Tvrdší“ (exaktnější)"))
levels(my[[likert.vars[i]]]) <-  c("Spíše ne", "Nevím", "Spíše ano")

# the variable "obor_mensi" is replaced by "obor_vetsi" with complementary values
my$obor_vetsi <- my$obor_mensi
levels(my$obor_vetsi) <- c("Rozhodně ano", "Spíše ano", "Nevím", "Spíše ne", "Rozhodně ne")
my$obor_vetsi <- ordered(my$obor_vetsi, levels=c("Rozhodně ne", "Spíše ne", "Nevím", "Spíše ano", "Rozhodně ano"))
explanatory.vars[explanatory.vars=="obor_mensi"] <- "obor_vetsi"
likert.vars[likert.vars=="obor_mensi"] <- "obor_vetsi" 
my <- subset(my,select=-obor_mensi)


# Nyní specifikujeme, které vysvětlující veličiny explanatory.vars se nemohou chápat pomocí Likertovy škály, ale jsou ordinální.
ordinal.vars <- c(
"vek",          
"kar_praxe",
"kar_titul",       
"kar_citace",      
"pomer_veda",
"pomer_zakladni",
"cinnost_matem",
"cinnost_model",
"cinnost_teori",
"cinnost_exper",
"cinnost_kateg",
"cinnost_vyvoj")
# u ordinálních veličin musíme správně nastavit uspořádání jejich hodnot
my[[ordinal.vars[1]]] <- ordered(my[[ordinal.vars[1]]], 
                                  levels = c("18–24","25–34","35–44","45–54","55–64","65 a více"))
my[[ordinal.vars[2]]] <- ordered(my[[ordinal.vars[2]]], 
                                  levels = c("Méně než 5 let","5–9 let","10–19 let","20 a více let"))
my[[ordinal.vars[3]]] <- ordered(my[[ordinal.vars[3]]], 
                                  levels = c("Bc.","Mgr./Ing.","CSc./Ph.D.","Docent","DrSc./DSc.","Profesor"))
my[[ordinal.vars[4]]] <- ordered(my[[ordinal.vars[4]]], 
                                  levels = c("Do 100","Do 500","Do 1000","Do 5000","Více"))
for (i in 5:6){
  my[!is.na(my[,ordinal.vars[i]]) & my[,ordinal.vars[i]]=="Nevím", ordinal.vars[5]] <- NA
  my[[ordinal.vars[i]]] <- ordered(my[[ordinal.vars[i]]], 
                                    levels = c("10%","20%","30%","40%","50%","60%","70%","80%","90%","100%"))
}
for (i in 7:12){
  my[[ordinal.vars[i]]] <- ordered(my[[ordinal.vars[i]]], 
                                    levels = c("1. místo","2. místo","3. místo","Tuto činnost nedělám"))
}

boolean.vars <- c(
  "kar_nejlepsi")

# Velmi složité najít uspořádání u veličiny kar_kategorie, která má 27 různých hodnot.
# Nebudeme ji používat.
complicated.vars <- c(
  "kar_kategorie")

# zbylé veličiny jsou kategorické 
categorical.vars <- setdiff(explanatory.vars,union(likert.vars,union(ordinal.vars,union(boolean.vars,complicated.vars))))

# For spec_hlavni - create shorter, cleaner names
my$spec_hlavni <- factor(my$spec_hlavni,
                         levels = c("Chemistry", "Biology", "Engineering", "Medicine", 
                                    "Mathematics", "Other specialization", "Materials Science", 
                                    "Physics", "Earth and Environmental Science", 
                                    "Astronomy and Planetary Science"),
                         labels = c("Chemistry", "Biology", "Engineering", "Medicine", 
                                    "Mathematics", "Other", "Materials", 
                                    "Physics", "Earth_Env", "Astronomy"))

# Set treatment contrasts also for ordered factors
# options(contrasts = c("contr.treatment", "contr.treatment"))
# This is the default:
options(contrasts = c("contr.treatment", "contr.poly"))


```
```{r}
library(corrplot)

# Prepare the data
vars <- union(setdiff(explanatory.vars,complicated.vars),"krize_rep")
my.vars.prepared <- my[,vars]
# my.vars.prepared <- my.vars

# Convert ordered factors to numeric (these are ordinal)
ordered_vars <- sapply(my.vars.prepared, is.ordered)
for(var in names(my.vars.prepared)[ordered_vars]) {
  my.vars.prepared[[var]] <- as.numeric(my.vars.prepared[[var]])
}

# Handle character variables
# kar_nejlepsi: binary variable (Ano/Ne)
if("kar_nejlepsi" %in% names(my.vars.prepared)) {
  my.vars.prepared$kar_nejlepsi <- as.numeric(as.factor(my.vars.prepared$kar_nejlepsi))
}

# gender: nominal categorical (3 levels)
if("gender" %in% names(my.vars.prepared)) {
  my.vars.prepared$gender <- as.numeric(as.factor(my.vars.prepared$gender))
}

# We exclude spec_hlavni from correlation computations since it's purely nominal with many categories (10 levels)
if("spec_hlavni" %in% names(my.vars.prepared)) {
  my.vars.for_cor <- my.vars.prepared[, !names(my.vars.prepared) %in% "spec_hlavni"]
}

print("Variables for correlation analysis:")
print(names(my.vars.for_cor))

# Compute correlations
cor_matrix <- cor(my.vars.for_cor, use = "pairwise.complete")

# Reorder the correlation matrix to put krize_rep first
desired_order <- c("krize_rep", setdiff(colnames(cor_matrix), "krize_rep"))
cor_matrix <- cor_matrix[desired_order, desired_order]

# Visualize the correlation matrix
pdf(file="correlation-matrix.pdf")
corrplot(cor_matrix, method = "color", type = "upper", 
         diag = FALSE, # Explicitly exclude diagonal
         tl.cex = 0.8, tl.col = "black", order="original")
dev.off()

```
The Fruchterman-Reingold algorithm visualizing correlation networks

```{r}
library(igraph)

# Set threshold
threshold <- 0.15

library(igraph)

# Create adjacency matrix with ABSOLUTE values
adj_matrix <- abs(cor_matrix)
adj_matrix[adj_matrix < threshold] <- 0
diag(adj_matrix) <- 0

# Create graph
cor_graph <- graph_from_adjacency_matrix(adj_matrix, 
                                        mode = "undirected", 
                                        weighted = TRUE)

# Get edge list and extract original correlation values
edge_list <- as_edgelist(cor_graph)
original_correlations <- sapply(1:nrow(edge_list), function(i) {
  v1 <- edge_list[i, 1]
  v2 <- edge_list[i, 2]
  return(cor_matrix[v1, v2])
})

# Add original correlations as edge attribute
E(cor_graph)$original_cor <- original_correlations

# Apply Fruchterman-Reingold layout
set.seed(123)
fr_layout <- layout_with_fr(cor_graph)

pdf(file="correlation-graph.pdf")
# Plot with edge width proportional to correlation strength
plot(cor_graph,
     layout = fr_layout,
     vertex.size = 12,
     vertex.color = ifelse(V(cor_graph)$name == "krize_rep", "gold", "lightsteelblue"),
     vertex.label.cex = 0.6,
     vertex.label.color = "black",
     edge.width = abs(E(cor_graph)$original_cor) * 8,  # Width based on absolute correlation
     edge.color = ifelse(E(cor_graph)$original_cor > 0, "red", "blue"),
     main = "Correlation Network (Fruchterman-Reingold Layout)")

legend("topright", 
       legend = c("Positive correlation", "Negative correlation", "Thicker = Stronger"),
       col = c("red", "blue", "black"),
       lty = c(1, 1, 1),
       lwd = c(2, 2, 4),
       cex = 0.8)
dev.off()

```



Correlation including Halm's correction

```{r}
# Load required libraries
library(corrplot)
library(igraph)
library(Hmisc)  # Alternative approach
library(psych)  # Another alternative

# First, check and clean the data
print("Checking data structure:")
print(str(my.vars.for_cor))
print("Class of object:")
print(class(my.vars.for_cor))

# Convert tibble to regular data frame
if("tbl_df" %in% class(my.vars.for_cor)) {
  cat("Converting tibble to data frame...\n")
  my.vars.for_cor <- as.data.frame(my.vars.for_cor)
} else if(is.list(my.vars.for_cor) && !is.data.frame(my.vars.for_cor)) {
  cat("Converting list to data frame...\n")
  my.vars.for_cor <- as.data.frame(my.vars.for_cor)
} else if(is.matrix(my.vars.for_cor)) {
  cat("Converting matrix to data frame...\n")
  my.vars.for_cor <- as.data.frame(my.vars.for_cor)
}

print("After conversion:")
print(class(my.vars.for_cor))
print("Data types:")
print(sapply(my.vars.for_cor, class))

# Keep only numeric columns
numeric_cols <- sapply(my.vars.for_cor, is.numeric)
if(sum(numeric_cols) == 0) {
  stop("No numeric columns found in the dataset!")
}

if(sum(!numeric_cols) > 0) {
  cat("Removing non-numeric columns:", paste(names(my.vars.for_cor)[!numeric_cols], collapse = ", "), "\n")
  my.vars.for_cor <- my.vars.for_cor[, numeric_cols, drop = FALSE]
}

# Remove columns with zero variance (constant values)
zero_var_cols <- sapply(my.vars.for_cor, function(x) var(x, na.rm = TRUE) == 0 | is.na(var(x, na.rm = TRUE)))
if(sum(zero_var_cols) > 0) {
  cat("Removing zero-variance columns:", paste(names(my.vars.for_cor)[zero_var_cols], collapse = ", "), "\n")
  my.vars.for_cor <- my.vars.for_cor[, !zero_var_cols, drop = FALSE]
}

# Check for minimum number of complete observations
complete_cases <- complete.cases(my.vars.for_cor)
cat("Complete cases:", sum(complete_cases), "out of", nrow(my.vars.for_cor), "\n")

if(sum(complete_cases) < 3) {
  stop("Insufficient complete cases for correlation analysis!")
}

cat("Final dataset dimensions:", dim(my.vars.for_cor), "\n")
cat("Final variables:", paste(names(my.vars.for_cor), collapse = ", "), "\n")

# Method 1: Using cor.test with apply (most flexible)
# Function to compute p-values for correlation matrix
compute_cor_pvalues <- function(data) {
  n_vars <- ncol(data)
  cor_matrix <- cor(data, use = "complete.obs")
  p_matrix <- matrix(NA, n_vars, n_vars)
  colnames(p_matrix) <- colnames(data)
  rownames(p_matrix) <- colnames(data)
  
  # Compute p-values for upper triangle
  for(i in 1:(n_vars-1)) {
    for(j in (i+1):n_vars) {
      # Extract vectors and ensure they're numeric
      x_vec <- as.numeric(data[,i])
      y_vec <- as.numeric(data[,j])
      
      # Remove NA pairs
      complete_pairs <- complete.cases(x_vec, y_vec)
      
      if(sum(complete_pairs) >= 3) {  # Need at least 3 observations
        test_result <- cor.test(x_vec[complete_pairs], y_vec[complete_pairs])
        p_matrix[i,j] <- test_result$p.value
        p_matrix[j,i] <- test_result$p.value  # Mirror to lower triangle
      } else {
        p_matrix[i,j] <- 1  # Set p-value to 1 for insufficient data
        p_matrix[j,i] <- 1
      }
    }
  }
  
  # Diagonal is always 1 (perfect correlation), p-value = 0
  diag(p_matrix) <- 0
  
  return(p_matrix)
}

# Compute correlation matrix and p-values
cor_matrix <- cor(my.vars.for_cor, use = "complete.obs")
p_matrix <- compute_cor_pvalues(my.vars.for_cor)

# Method 2: Using Hmisc::rcorr (recommended - more robust)
# Convert to matrix for rcorr
data_matrix <- as.matrix(my.vars.for_cor)

# Check if conversion worked
if(!is.numeric(data_matrix)) {
  cat("Matrix conversion failed. Attempting column-by-column conversion...\n")
  for(i in 1:ncol(my.vars.for_cor)) {
    my.vars.for_cor[,i] <- as.numeric(as.character(my.vars.for_cor[,i]))
  }
  data_matrix <- as.matrix(my.vars.for_cor)
}

result <- rcorr(data_matrix)
cor_matrix <- result$r
p_matrix <- result$P

# Apply Holm correction to p-values
# Extract upper triangle p-values (excluding diagonal)
upper_tri_indices <- which(upper.tri(p_matrix), arr.ind = TRUE)
p_values_vector <- p_matrix[upper.tri(p_matrix)]

# Apply Holm correction
p_adjusted <- p.adjust(p_values_vector, method = "holm")

# Create adjusted p-value matrix
p_matrix_adjusted <- p_matrix
p_matrix_adjusted[upper.tri(p_matrix_adjusted)] <- p_adjusted
p_matrix_adjusted[lower.tri(p_matrix_adjusted)] <- t(p_matrix_adjusted)[lower.tri(p_matrix_adjusted)]

# Set significance threshold
alpha <- 0.05
# alpha <- 0.2

# Create significance mask
sig_mask <- p_matrix_adjusted < alpha

# Filter correlation matrix based on significance
cor_matrix_filtered <- cor_matrix * sig_mask

# Optional: Set a minimum correlation threshold for network edges
cor_threshold <- 0.15
# cor_threshold <- 0.3  # Adjust as needed
cor_matrix_filtered[abs(cor_matrix_filtered) < cor_threshold] <- 0

# Create network graph for Fruchterman-Reingold algorithm
# Convert correlation matrix to adjacency matrix (absolute values for undirected graph)
adj_matrix <- abs(cor_matrix_filtered)

# Remove diagonal (self-loops)
diag(adj_matrix) <- 0

# Create igraph object (weights will be automatically set from the adjacency matrix)
network <- graph_from_adjacency_matrix(adj_matrix, 
                                     mode = "undirected", 
                                     weighted = TRUE, 
                                     diag = FALSE)

# The weights are already set correctly by graph_from_adjacency_matrix
# But let's verify and set edge colors properly

# Get the edge list to properly assign colors
edge_list <- as_edgelist(network, names = FALSE)
n_edges <- nrow(edge_list)

# Create vectors to store edge attributes
edge_weights <- numeric(n_edges)
edge_colors <- character(n_edges)

# For each edge, get the corresponding correlation value
for(i in 1:n_edges) {
  row_idx <- edge_list[i, 1]
  col_idx <- edge_list[i, 2]
  
  # Get the original correlation value (with sign)
  cor_value <- cor_matrix_filtered[row_idx, col_idx]
  
  # Set weight as absolute correlation
  edge_weights[i] <- abs(cor_value)
  
  # Set color based on sign
  edge_colors[i] <- ifelse(cor_value > 0, "red", "blue")
}

# Assign the attributes
E(network)$weight <- edge_weights
E(network)$color <- edge_colors

# Apply Fruchterman-Reingold layout
set.seed(123)  # For reproducible layout
fr_layout <- layout_with_fr(network)

pdf(file="correlation-graph-Holms-correction.pdf")
# Plot the network
plot(network, 
     layout = fr_layout,
     vertex.size = 12,
     vertex.color = ifelse(V(cor_graph)$name == "krize_rep", "gold", "lightsteelblue"),
     vertex.label.cex = 0.6,
     vertex.label.color = "black",
     edge.width = E(network)$weight * 8,  # Scale edge width by correlation strength
     edge.color = E(network)$color,
     main = "Correlation Network (Holm-corrected, Fruchterman-Reingold Layout)")

# Add legend
legend("topleft", 
       legend = c("Positive correlation", "Negative correlation", "Thicker = Stronger"),
       col = c("red", "blue", "black"),
       lty = c(1, 1, 1),
       lwd = c(2, 2, 4),
       cex = 0.8)
dev.off()

# Print summary statistics
cat("Original correlations - significant pairs (before correction):", 
    sum(p_matrix < alpha, na.rm = TRUE) / 2, "\n")
cat("Holm-corrected - significant pairs:", 
    sum(p_matrix_adjusted < alpha, na.rm = TRUE) / 2, "\n")
cat("Network edges (after filtering):", ecount(network), "\n")
cat("Network nodes:", vcount(network), "\n")

# Optional: Save results
# write.csv(cor_matrix, "correlation_matrix.csv")
# write.csv(p_matrix_adjusted, "p_values_holm_corrected.csv")

# View first few rows of results
print("Correlation matrix (first 5x5):")
print(round(cor_matrix[1:min(5,nrow(cor_matrix)), 1:min(5,ncol(cor_matrix))], 3))

print("Holm-corrected p-values (first 5x5):")
print(round(p_matrix_adjusted[1:min(5,nrow(p_matrix_adjusted)), 1:min(5,ncol(p_matrix_adjusted))], 4))

```



```{r}
library(bnlearn)
my.vars.bn <- my.vars.for_cor
# if we want to consider also this variable with 10 values
my.vars.bn <- cbind(my.vars.for_cor, my_clean$spec_hlavni)
colnames(my.vars.bn) <- c(colnames(my.vars.bn)[-length(colnames(my.vars.bn))],"spec_hlavni")

# Convert all columns to factors at once
my.vars.bn.discrete <- as.data.frame(lapply(my.vars.bn, as.factor))

# Check the structure to confirm conversion
str(my.vars.bn.discrete)

library(dplyr)
# Filter to only existing variables
missing_pct <- my.vars.bn.discrete %>%
  dplyr::select(all_of(names(my.vars.bn.discrete ))) %>%
  summarise_all(~mean(is.na(.)) * 100)
# Keep only variables with <10% missing
good_vars <- names(missing_pct)[missing_pct < 10]
# These are variables that are frequently missing
print(setdiff(all.predictors,good_vars))

# Remove columns with too many missing values first, then rows
my.vars.bn.clean <- my.vars.bn.discrete[, good_vars]
my.vars.bn.clean <- na.omit(my.vars.bn.clean)

# Now we can learn the Bayesian network structure
bn_structure <- bnlearn::hc(my.vars.bn.clean, score = "aic")
# Number of edges for different scoring functions
#   BIC (default): 28 edges
#   AIC: 40 edges
#   Log-likelihood: 288 edges

# Plot the learned network
plot(bn_structure)
# amat(bn_structure)
# To get more information about the learned network:
arcs(bn_structure)  # Shows the directed edges
# nodes(bn_structure) # Shows the nodes
score(bn_structure, my.vars.bn.clean) # BIC score

# Other structure learning algorithms:
bn_structure_tabu <- tabu(my.vars.bn.clean)
plot(bn_structure_tabu)
#
bn_structure_mmhc <- mmhc(my.vars.bn.clean)
plot(bn_structure_mmhc)


```

Visualization of learned Bayesian networks

```{r}

# Using DiagrammeR (modern alternative)
# Install if needed: install.packages("DiagrammeR")
if (requireNamespace("DiagrammeR", quietly = TRUE)) {
  library(DiagrammeR)
  
  # Convert bnlearn network (including CPDAG) to DOT format
  create_dot_bn <- function(bn_object, title = "Bayesian Network") {
    nodes_list <- nodes(bn_object)
    edges_list <- arcs(bn_object)
    
    # Identify undirected edges (edges that appear in both directions)
    directed_edges <- data.frame()
    undirected_edges <- data.frame()
    processed_pairs <- character(0)
    
    if(nrow(edges_list) > 0) {
      for(i in 1:nrow(edges_list)) {
        from <- edges_list[i,1]
        to <- edges_list[i,2]
        pair_key <- paste(sort(c(from, to)), collapse = "_")
        
        if(!pair_key %in% processed_pairs) {
          # Check if reverse edge exists
          reverse_exists <- any(edges_list[,1] == to & edges_list[,2] == from)
          
          if(reverse_exists) {
            # Undirected edge - add only once
            undirected_edges <- rbind(undirected_edges, data.frame(from = from, to = to))
          } else {
            # Directed edge
            directed_edges <- rbind(directed_edges, data.frame(from = from, to = to))
          }
          processed_pairs <- c(processed_pairs, pair_key)
        }
      }
    }
    
    # Start DOT string
    dot_string <- 'digraph BN {
  graph [layout="dot", rankdir="TB", bgcolor="white"];
  node [shape="ellipse", style="filled", fillcolor="lightsteelblue", 
        color="navy", fontcolor="black", fontname="Arial", fontsize="12"];
  
'
    
    # Add nodes with labels
    for(node in nodes_list) {
      # Keep original node names with underscores
      dot_string <- paste0(dot_string, '  "', node, '" [label="', node, '"];\n')
    }
    
    dot_string <- paste0(dot_string, '\n')
    
    # Add directed edges
    if(nrow(directed_edges) > 0) {
      dot_string <- paste0(dot_string, '  // Directed edges\n')
      for(i in 1:nrow(directed_edges)) {
        dot_string <- paste0(dot_string, '  "', directed_edges[i,1], '" -> "', 
                           directed_edges[i,2], '" [arrowhead="vee", color="darkblue", penwidth="2"];\n')
      }
    }
    
    # Add undirected edges
    if(nrow(undirected_edges) > 0) {
      dot_string <- paste0(dot_string, '\n  // Undirected edges\n')
      for(i in 1:nrow(undirected_edges)) {
        dot_string <- paste0(dot_string, '  "', undirected_edges[i,1], '" -> "', 
                           undirected_edges[i,2], '" [arrowhead="none", arrowtail="none", dir="none", color="darkblue", penwidth="2"];\n')
      }
    }
    
    dot_string <- paste0(dot_string, '}')
    return(dot_string)
  }
  
  # Create and render the graph
  dot_graph <- create_dot_bn(cpdag(bn_structure))
  print(grViz(dot_graph))
  
} else {
  cat("DiagrammeR not available. Install with: install.packages('DiagrammeR')\n")
}

cpdag_structure <- cpdag(bn_structure)
dot_content <- create_dot_bn(cpdag_structure)
writeLines(dot_content, "network.dot")
system('"C:/Program Files/Graphviz/bin/dot.exe" -Tpdf network.dot -o cpdag-of-bayesian-network.pdf')


```


Multinomial Regression using nnet package 

```{r}
# learning multinomial regression model
library(nnet)
# Combine all predictor variables
all.predictors <- c(likert.vars, ordinal.vars, boolean.vars, categorical.vars)
# Create formula using reformulate
formula <- reformulate(termlabels = all.predictors, response = dependent.var)

# Fit multinomial logistic regression
# decay = 0.01
# decay = 1.8
model <- multinom(formula, data = my, decay = 1.8, # L2 penalty parameter 
                  maxit = 1000 # Increase iterations
                  )
```

The problem of this model is that it uses very few respondents for learning

```{r}
model_frame_used <- model.frame(model)
# Now we can see it uses only 260 data vectors, which is much less than nrow(my), which is 878.
print(nrow(model_frame_used))
# multinom doesn't have an option for variable-specific sample sizes. 
# This is a fundamental limitation of how regression models work. 
# Multiple imputation is the proper statistical solution to use all available information while maintaining valid inference, use e.g. mice ...
```

We remove variables with too much missingness rather than rows

```{r}
library(dplyr)
# Filter to only existing variables
valid.predictors <- all.predictors[all.predictors %in% names(my)]
missing_pct <- my %>%
  dplyr::select(all_of(valid.predictors)) %>%
  summarise_all(~mean(is.na(.)) * 100)
# Keep only variables with <10% missing
good_vars <- names(missing_pct)[missing_pct < 25]
# These are variables that are frequently missing
print(setdiff(all.predictors,good_vars))
```

Now we learn model with variables that do not have high missingness

```{r}
formula_reduced <- reformulate(termlabels = good_vars, response = dependent.var)
# Fit multinomial logistic regression
# decay = 0.01
# decay = 1.8
model <- multinom(formula_reduced, data = my, decay = 1.8, # L2 penalty parameter 
                  maxit = 1000 # Increase iterations
                  )
model_frame_used <- model.frame(model)
# Now we can see it uses only 747 data vectors, which is not so much less than nrow(my), which is 878.
print(nrow(model_frame_used))
BIC(model)
```
```{r}
actual_y <- model.response(model_frame_used)

  # View results
  # summary(model)
  # Get coefficients
  # coef(model)
  # Predict probabilities
  predicted_probs <- predict(model, type = "probs")
  # Make predictions
  predictions <- predict(model, type = "class")
  # compare predictions with actual values in data
  library(caret)
  cm <- confusionMatrix(data = predictions, reference = as.factor(actual_y))
  print(cm)
  print.cfM(cm$table, file="cfM-mlr.pdf")
```

Quick systematic approach to find significance of variables using the model likelihood ratio (LR)
Note: 
for decay=1.8 found by caret 10F CV no variable is significant 
for decay=0 two variables are significant: obor_rychly and vek

```{r}

# Enhanced version with significance stars
test_variable_significance_enhanced <- function(full_model, data, decay=1.8) {
  # Get all variables used in the model
  model_vars <- all.vars(formula(full_model))
  complete_data <- data[complete.cases(data[, model_vars]), ]
  
  print(paste("Using", nrow(complete_data), "complete observations out of", nrow(data)))
  
  # Refit full model on complete data
  full_model_complete <- multinom(formula(full_model), data = complete_data, trace = FALSE, 
                                  decay = decay, # L2 penalty parameter 
                                  maxit = 1000 # Increase iterations
  )
  full_loglik <- logLik(full_model_complete)
  
  # Get predictor variables (excluding response)
  response_var <- all.vars(formula(full_model))[1]
  predictor_vars <- setdiff(all.vars(formula(full_model)), response_var)
  
  results <- data.frame(
    Variable = character(),
    LR_Chisq = numeric(),
    Df = numeric(),
    P_value = numeric(),
    Significance = character(),
    stringsAsFactors = FALSE
  )
  
  for (var in predictor_vars) {
    cat("Testing variable:", var, "\n")
    
    # Fit reduced model on the SAME complete data
    reduced_formula <- update(formula(full_model), paste("~ . -", var))
    reduced_model <- multinom(reduced_formula, data = complete_data, trace = FALSE,
                              decay = decay, # L2 penalty parameter 
                              maxit = 1000 # Increase iterations
    )
    reduced_loglik <- logLik(reduced_model)
    
    # Calculate LR statistic
    lr_stat <- -2 * (as.numeric(reduced_loglik) - as.numeric(full_loglik))
    df_diff <- attr(full_loglik, "df") - attr(reduced_loglik, "df")
    p_value <- pchisq(lr_stat, df = df_diff, lower.tail = FALSE)
    
    # Add significance stars
    sig_stars <- ifelse(p_value < 0.001, "***",
                 ifelse(p_value < 0.01, "**",
                 ifelse(p_value < 0.05, "*",
                 ifelse(p_value < 0.1, ".", ""))))
    
    results <- rbind(results, data.frame(
      Variable = var,
      LR_Chisq = round(lr_stat, 3),
      Df = df_diff,
      P_value = round(p_value, 5),
      Significance = sig_stars
    ))
  }
  
  # Order by p-value (ascending)
  results <- results[order(results$P_value), ]
  
  # Reset row names
  rownames(results) <- NULL
  
  return(results)
}

print("Output with significance indicators:")
enhanced_results <- test_variable_significance_enhanced(model, my, decay=0)
print(enhanced_results)
print("Significance codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1")

```

Plotting the influence of explanatory variables as represented in the learned model

```{r}
library(effects)

# Create effect plot for variable "vek"
age_effect <- Effect("vek", model)
plot(age_effect, style = "stacked")  # or "lines" for separate lines


# Create effect plot for variable "kar_praxe"
age_effect <- Effect("kar_praxe", model)
plot(age_effect, style = "stacked")  # or "lines" for separate lines

# Create effect plot for variable "pomer_zakladni"
basic_research_effect <- Effect("pomer_zakladni", model)
plot(basic_research_effect, style = "stacked")  # or "lines" for separate lines
```

Tuning model parameters using 10-Fold Cross-Validation

```{r}
library(caret)
set.seed(123)
train_control<- trainControl(method="cv", number=10, savePredictions = TRUE)
tune_grid <- expand.grid(decay = seq(0, 2, by = 0.2))
model_10f <- train(formula_reduced, 
                       data=my, 
                       trControl=train_control, 
                       method="multinom", 
                       na.action = na.omit, 
                       tuneGrid = tune_grid)
# Look at the CV performance
print(model_10f)
model_10f$results

# Look at the final model
summary(model_10f$finalModel)

# Variable importance from the CV model
# varImp(olr_model_10f)
# plot(varImp(olr_model_10f), top = 10)

# Get predictions from finalModel (747 records)
final_predictions <- predict(model_10f$finalModel)
# The finalModel used 747 out of 806 records from trainingData
# We need to figure out which 747 records were actually used
training_data <- model_10f$trainingData
# Check if there are any additional NAs that were removed
complete_rows <- complete.cases(training_data[, -ncol(training_data)])  # Exclude .outcome column
actual_outcomes <- training_data$.outcome[complete_rows]

# Create confusion matrix
cm <- confusionMatrix(final_predictions, actual_outcomes)
print(cm)
print.cfM(cm$table, file="cfM-mlr-10f.pdf")
```
```{r}
source("../../CBSllm/CV-functions-v18.R")

# Run cross-validation
results <- greedy_forward_selection(
  data = my[,c(good_vars,dependent.var)],
  response_var = dependent.var,
  decay_val = c(0.01, 0.1, 1, 2, 5, 10, 20),
)

```

Compute correlations between predictions and actual outcomes

```{r}
spearman_cor <- cor(as.numeric(final_predictions), as.numeric(actual_outcomes), 
                    method = "spearman")
print(paste("Spearman correlation:", round(spearman_cor, 3)))

pearson_cor <- cor(as.numeric(final_predictions), as.numeric(actual_outcomes), 
                    method = "pearson")
print(paste("Pearson  correlation:", round(pearson_cor, 3)))

```

Model parameters

```{r}
multinom_model <- model_10f

# Method 1: Basic coefficient extraction
# Extract the final model from caret train object
final_model <- multinom_model$finalModel

# Get raw coefficients
raw_coefficients <- coef(final_model)
cat("Raw coefficient matrix:\n")
print(raw_coefficients)
cat("\nDimensions:", dim(raw_coefficients), "(rows = classes, columns = predictors)\n")

# Method 2: Clean coefficient extraction with proper names
# Get coefficient names
coef_names <- colnames(raw_coefficients)
class_names <- rownames(raw_coefficients)

cat("\nCoefficient names:", paste(coef_names, collapse = ", "), "\n")
cat("Class names:", paste(class_names, collapse = ", "), "\n")

# Method 3: Convert to data frame for easier handling
coef_df <- as.data.frame(raw_coefficients)
coef_df$Class <- rownames(raw_coefficients)
coef_df <- coef_df[, c("Class", setdiff(names(coef_df), "Class"))]

cat("\nCoefficients as data frame:\n")
print(coef_df)

# Method 4: Long format for easier analysis/plotting
library(tidyr)
library(dplyr)

coef_long <- coef_df %>%
  pivot_longer(cols = -Class, names_to = "Predictor", values_to = "Coefficient") %>%
  arrange(Class, Predictor)

cat("\nCoefficients in long format (first 20 rows):\n")
print(head(coef_long, 20))

# Method 5: Get standard errors and compute significance
model_summary <- summary(final_model)
standard_errors <- model_summary$standard.errors
z_values <- model_summary$coefficients / standard_errors
p_values <- 2 * (1 - pnorm(abs(z_values)))

# Create comprehensive coefficient table
coef_detailed <- data.frame(
  Class = rep(rownames(raw_coefficients), each = ncol(raw_coefficients)),
  Predictor = rep(colnames(raw_coefficients), nrow(raw_coefficients)),
  Coefficient = as.vector(t(raw_coefficients)),
  Std_Error = as.vector(t(standard_errors)),
  Z_value = as.vector(t(z_values)),
  P_value = as.vector(t(p_values))
)

# Add significance stars
coef_detailed$Significance <- cut(coef_detailed$P_value,
                                 breaks = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                                 labels = c("***", "**", "*", ".", ""),
                                 include.lowest = TRUE)

cat("\nDetailed coefficient table (first 20 rows):\n")
print(head(coef_detailed, 20))

# Method 6: Calculate odds ratios and confidence intervals
odds_ratios <- exp(raw_coefficients)
cat("\nOdds ratios:\n")
print(round(odds_ratios, 3))

# Try to get confidence intervals (may not always work)
tryCatch({
  conf_intervals <- confint(final_model)
  odds_ratio_ci <- exp(conf_intervals)
  cat("\nConfidence intervals for odds ratios:\n")
  print(round(odds_ratio_ci, 3))
}, error = function(e) {
  cat("Could not compute confidence intervals:", e$message, "\n")
  
  # Alternative: approximate CI using standard errors
  cat("Computing approximate 95% confidence intervals...\n")
  ci_lower <- raw_coefficients - 1.96 * standard_errors
  ci_upper <- raw_coefficients + 1.96 * standard_errors
  
  # Convert to odds ratios
  or_ci_lower <- exp(ci_lower)
  or_ci_upper <- exp(ci_upper)
  
  cat("Odds ratio confidence intervals (approximate):\n")
  cat("Lower bounds:\n")
  print(round(or_ci_lower, 3))
  cat("Upper bounds:\n")  
  print(round(or_ci_upper, 3))
})

# Method 7: Visualize coefficients with improved text sizing and renamed classes
library(ggplot2)

# First, let's see what the current class names are
current_classes <- unique(coef_long$Class)
cat("Current class names:", paste(current_classes, collapse = ", "), "\n")

class_mapping <- c(
  "Ano, nevýznamná krize" = "Yes, minor crisis", 
  "Ne, žádná krize není" = "No, there is no crisis", 
  "Nevím" = "I don't know"
)
# Apply the mapping to data
coef_long_renamed <- coef_long %>%
  mutate(Class_Original = Class,  # Keep original for reference
         Class = ifelse(Class %in% names(class_mapping), 
                       class_mapping[Class], 
                       Class))  # Use mapping if available, otherwise keep original
# Update other data frames as well
coef_detailed_renamed <- coef_detailed %>%
  mutate(Class_Original = Class,
         Class = ifelse(Class %in% names(class_mapping), 
                       class_mapping[Class], 
                       Class))
# Show the mapping
cat("Class name mapping:\n")
mapping_df <- data.frame(
  Original = names(class_mapping),
  Renamed = class_mapping
)
print(mapping_df)

pdf(file="mlr-model.pdf")
# Plot coefficients by class
p1 <- ggplot(coef_long_renamed, aes(x = Coefficient, y = reorder(Predictor, abs(Coefficient)), 
                           color = Class)) +
  geom_point(size = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~Class, scales = "free_x") +
  labs(title = "Multinomial Coefficients by Class",
       x = "Coefficient Value",
       y = "Predictors") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.y = element_text(size = 6))
print(p1)
dev.off()

# Heatmap of coefficients
coef_matrix_plot <- coef_long %>%
  mutate(Predictor = factor(Predictor, levels = unique(Predictor)))

p2 <- ggplot(coef_matrix_plot, aes(x = Class, y = Predictor, fill = Coefficient)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "Coefficient Heatmap",
       subtitle = "Blue = negative, Red = positive coefficients") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p2)

# Method 8: Extract coefficients for specific predictors or classes
# For specific predictor across all classes
get_predictor_coefs <- function(predictor_name) {
  if(predictor_name %in% coef_names) {
    result <- raw_coefficients[, predictor_name, drop = FALSE]
    cat(paste("Coefficients for", predictor_name, ":\n"))
    print(result)
    return(result)
  } else {
    cat("Predictor not found. Available predictors:\n")
    print(coef_names)
  }
}

# For specific class across all predictors  
get_class_coefs <- function(class_name) {
  if(class_name %in% class_names) {
    result <- raw_coefficients[class_name, , drop = FALSE]
    cat(paste("Coefficients for class", class_name, ":\n"))
    print(result)
    return(result)
  } else {
    cat("Class not found. Available classes:\n")
    print(class_names)
  }
}

# Example usage (uncomment to use):
# get_predictor_coefs("your_predictor_name")
# get_class_coefs("your_class_name")

# Method 9: Export coefficients
# Save to CSV
write.csv(coef_detailed, "multinomial_coefficients_detailed.csv", row.names = FALSE)
write.csv(coef_df, "multinomial_coefficients_matrix.csv", row.names = FALSE)

cat("\nCoefficients saved to CSV files:\n")
cat("- multinomial_coefficients_detailed.csv (long format with statistics)\n")
cat("- multinomial_coefficients_matrix.csv (matrix format)\n")

# Method 10: Summary interpretation
cat("\n=== COEFFICIENT INTERPRETATION ===\n")
cat("For multinomial logistic regression:\n")
cat("1. Coefficients are log-odds relative to the reference class\n")
cat("2. Positive coefficient = increases odds of that class vs reference\n")
cat("3. Negative coefficient = decreases odds of that class vs reference\n")
cat("4. Magnitude indicates strength of effect\n")
cat("5. For ordered factors: coefficients show trend across levels\n\n")

# Find largest coefficients
largest_coefs <- coef_detailed %>%
  arrange(desc(abs(Coefficient))) %>%
  head(10)

cat("Top 10 largest coefficients (by absolute value):\n")
print(largest_coefs[, c("Class", "Predictor", "Coefficient", "P_value", "Significance")])

# Find most significant coefficients
most_significant <- coef_detailed %>%
  arrange(P_value) %>%
  head(10)

cat("\nTop 10 most significant coefficients:\n")
print(most_significant[, c("Class", "Predictor", "Coefficient", "P_value", "Significance")])
```


The variables imporatance analysis

```{r}
# Run the complete analysis
analysis <- analyze_multinom_importance(model_10f$finalModel)

# Display the summary plot (aggregated importance)
print(analysis$summary_plot)

# Display the detailed coefficient plot
print(analysis$detailed_plot)

# View the importance summary table
View(analysis$importance)

# View all coefficients and p-values
View(analysis$results)
```

Alternative approach to evaluate variables importance

```{r}
# Load required libraries
library(caret)
library(ggplot2)
library(dplyr)
library(viridis)
library(gridExtra)

multinom_model <- model_10f

# Extract variable importance from your caret model
var_imp <- varImp(multinom_model)

# Print the structure to understand what we're working with
cat("Structure of variable importance object:\n")
print(str(var_imp))

# Extract the importance scores
imp_scores <- var_imp$importance
cat("\nVariable importance scores:\n")
print(imp_scores)

# INTERPRETATION:
# For multinomial logistic regression in caret:
# - Values represent the absolute value of the t-statistic for each predictor
# - Higher values = more important for distinguishing between outcome classes
# - Values are typically scaled to 0-100 (100 = most important variable)
# - For multiclass outcomes, you get importance for each class comparison

# Method 1: Basic caret plot (built-in visualization)
cat("\nBasic caret variable importance plot:\n")
plot(var_imp, main = "Variable Importance from Caret Model")

# Method 2: Enhanced ggplot visualization
# Prepare data for ggplot
if(ncol(imp_scores) > 1) {
  # For multiclass - reshape data for plotting
  imp_long <- imp_scores %>%
    tibble::rownames_to_column("Variable") %>%
    tidyr::gather(key = "Class", value = "Importance", -Variable)
  
  # Plot 1: Heatmap for multiclass importance
  p1 <- ggplot(imp_long, aes(x = Class, y = reorder(Variable, Importance), fill = Importance)) +
    geom_tile() +
    scale_fill_viridis_c(name = "Importance") +
    labs(title = "Variable Importance by Class",
         subtitle = "Higher values = more important for distinguishing that class",
         x = "Outcome Class", 
         y = "Variables") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Plot 2: Overall importance (max across classes)
  imp_overall <- imp_scores %>%
    tibble::rownames_to_column("Variable") %>%
    mutate(Max_Importance = apply(select(., -Variable), 1, max),
           Mean_Importance = apply(select(., -Variable), 1, mean)) %>%
    arrange(desc(Max_Importance))
  
  p2 <- ggplot(imp_overall, aes(x = reorder(Variable, Max_Importance), y = Max_Importance)) +
    geom_col(fill = "steelblue", alpha = 0.7) +
    geom_text(aes(label = round(Max_Importance, 1)), 
              hjust = -0.1, size = 3) +
    coord_flip() +
    labs(title = "Overall Variable Importance",
         subtitle = "Maximum importance across all classes",
         x = "Variables", 
         y = "Importance Score") +
    theme_minimal()
  
  # Plot 3: Class-specific importance (faceted)
  p3 <- ggplot(imp_long, aes(x = reorder(Variable, Importance), y = Importance, fill = Class)) +
    geom_col() +
    coord_flip() +
    facet_wrap(~Class, scales = "free_x") +
    scale_fill_viridis_d(name = "Class") +
    labs(title = "Variable Importance by Class (Detailed)",
         x = "Variables", 
         y = "Importance Score") +
    theme_minimal() +
    theme(legend.position = "bottom")
  
} else {
  # For binary classification or single importance score
  imp_df <- data.frame(
    Variable = rownames(imp_scores),
    Importance = imp_scores[,1]
  ) %>%
    arrange(desc(Importance))
  
  p1 <- ggplot(imp_df, aes(x = reorder(Variable, Importance), y = Importance)) +
    geom_col(fill = "steelblue", alpha = 0.7) +
    geom_text(aes(label = round(Importance, 1)), 
              hjust = -0.1, size = 3) +
    coord_flip() +
    labs(title = "Variable Importance",
         subtitle = "Higher values indicate more predictive variables",
         x = "Variables", 
         y = "Importance Score") +
    theme_minimal()
  
  p2 <- p1  # Same plot for single class
  p3 <- p1  # Same plot for single class
}

# Display the plots
print(p1)
print(p2)
if(exists("p3")) print(p3)

# Method 3: Interactive plot using plotly (if available)
if(requireNamespace("plotly", quietly = TRUE)) {
  library(plotly)
  
  if(exists("imp_overall")) {
    p_interactive <- plot_ly(imp_overall, 
                           x = ~Max_Importance, 
                           y = ~reorder(Variable, Max_Importance),
                           type = 'bar',
                           orientation = 'h',
                           text = ~paste("Variable:", Variable, 
                                       "<br>Max Importance:", round(Max_Importance, 2),
                                       "<br>Mean Importance:", round(Mean_Importance, 2)),
                           textposition = 'none',
                           hovertemplate = "%{text}<extra></extra>") %>%
      layout(title = "Interactive Variable Importance",
             xaxis = list(title = "Importance Score"),
             yaxis = list(title = "Variables"))
    
    print(p_interactive)
  }
}

# Method 4: Comparison with other importance measures
cat("\n=== INTERPRETING THE NUMBERS ===\n")
cat("For multinomial logistic regression in caret:\n")
cat("1. Scores typically range from 0-100 (scaled)\n")
cat("2. 100 = most important variable in the model\n")
cat("3. Values represent |t-statistic| for each coefficient\n")
cat("4. Higher values = better at distinguishing between classes\n")
cat("5. For ordered factors: importance shows discriminatory power\n\n")

# Create interpretation table
if(exists("imp_overall")) {
  interpretation <- imp_overall %>%
    mutate(
      Interpretation = case_when(
        Max_Importance >= 80 ~ "Very High - Strong predictor",
        Max_Importance >= 60 ~ "High - Important predictor", 
        Max_Importance >= 40 ~ "Moderate - Useful predictor",
        Max_Importance >= 20 ~ "Low - Weak predictor",
        TRUE ~ "Very Low - Minimal impact"
      )
    ) %>%
    select(Variable, Max_Importance, Interpretation)
  
  cat("Variable Importance Interpretation:\n")
  print(interpretation)
} else if(exists("imp_df")) {
  interpretation <- imp_df %>%
    mutate(
      Interpretation = case_when(
        Importance >= 80 ~ "Very High - Strong predictor",
        Importance >= 60 ~ "High - Important predictor", 
        Importance >= 40 ~ "Moderate - Useful predictor",
        Importance >= 20 ~ "Low - Weak predictor",
        TRUE ~ "Very Low - Minimal impact"
      )
    )
  
  cat("Variable Importance Interpretation:\n")
  print(interpretation)
}

# Method 5: Compare with permutation importance from earlier
if(exists("perm_importance")) {
  comparison_df <- data.frame(
    Variable = names(perm_importance),
    Caret_Importance = if(exists("imp_overall")) imp_overall$Max_Importance[match(names(perm_importance), imp_overall$Variable)] else imp_df$Importance[match(names(perm_importance), imp_df$Variable)],
    Permutation_Importance = perm_importance * 100  # Scale to match caret
  )
  
  p_comparison <- ggplot(comparison_df, aes(x = Caret_Importance, y = Permutation_Importance)) +
    geom_point(size = 3, alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE, color = "red") +
    geom_text(aes(label = Variable), nudge_y = 2, size = 3) +
    labs(title = "Caret vs Permutation Importance",
         subtitle = "Agreement between methods indicates robust importance",
         x = "Caret Variable Importance",
         y = "Permutation Importance (scaled)") +
    theme_minimal()
  
  print(p_comparison)
  
  # Calculate correlation
  cor_imp <- cor(comparison_df$Caret_Importance, comparison_df$Permutation_Importance, use = "complete.obs")
  cat(paste("\nCorrelation between caret and permutation importance:", round(cor_imp, 3), "\n"))
}

cat("\n=== RECOMMENDATIONS FOR ORDERED FACTORS ===\n")
cat("1. Focus on variables with importance > 40 for practical significance\n")
cat("2. For ordered factors, high importance suggests strong trend across levels\n")
cat("3. Consider the class-specific importance for understanding which outcomes are best predicted\n")
cat("4. Cross-validate importance with other methods (LR tests, stepwise selection)\n")
cat("5. Variables important across multiple methods are most reliable\n")
```

Ordinal logistic regression using the polr 

```{r}
library(MASS)
olr_model <- polr(data=my, formula = formula_reduced, method = "logistic", Hess=TRUE)
BIC(olr_model)
```

Computing p-values for all variables of the model

```{r}
# Quick systematic approach to find significant variables:
library(car)
print("Significance of all variables:")
Anova(olr_model, type = "II")

if (FALSE) {
  # This works only for data with no missing values, which is NOT our case.
  print("\nBackward selection to find parsimonious model:")
  final_model <- step(olr_model, direction = "backward", test = "Chisq")
  summary(final_model)
}
```

Computing p-values independently for all but contrast value of all variables

```{r}
## store the coefficients table without the zeta coefficients
ctable <- coef(summary(olr_model))
ctable <- ctable[1:(nrow(ctable) - 3), ]
## calculate and store p values
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
## combined table
ctable <- cbind(ctable, "p value" = p)
ctable <- ctable[order(ctable[,"p value"]),]
nms <- rownames(ctable[ctable[,"p value"] <= 0.05,])

# variables with at least one significant coefficient for its values
unique(substr(nms,1,nchar(nms)-2))
```

Example of testing of a variable's significance by removing it from the model

```{r}
test.var <- "obor_dotovany"
# Get all terms from your original formula
all_terms <- attr(terms(formula_reduced), "term.labels")
# Remove the term you don't want
reduced_terms <- all_terms[!grepl(test.var, all_terms)]
# Rebuild formula
response_var <- all.vars(formula_reduced)[1]  # assumes first variable is response
reduced_formula <- as.formula(paste(response_var, "~", paste(reduced_terms, collapse = " + ")))
reduced_model <- polr(reduced_formula, data = my)
anova.res <- anova(reduced_model, olr_model)
# if the value is <= 0.05 the model with test.var is better
anova.res$`Pr(Chi)`[2]
```

```{r}
model_frame_used <- model.frame(olr_model)
actual_y <- model.response(model_frame_used)

  # View results
  # summary(model)
  # Get coefficients
  # coef(model)
  # Predict probabilities
  predicted_probs <- predict(olr_model, type = "probs")
  # Make predictions
  predictions <- predict(olr_model, type = "class")
  # compare predictions with actual values in data
  library(caret)
  cm <- confusionMatrix(data = predictions, reference = as.factor(actual_y))
  print(cm)
  print.cfM(cm$table, file="cfM-olr.pdf")
```

Caret package 10-fold cross-validation

```{r}
library(caret)
set.seed(123)
train_control<- trainControl(method="cv", number=10, savePredictions = TRUE)
olr_model_10f <- train(formula_reduced, data=my, trControl=train_control, method="polr", 
                       na.action = na.omit, tuneGrid = data.frame(method = "logistic"))
# Look at the CV performance
print(olr_model_10f)
olr_model_10f$results

# Look at the final model
summary(olr_model_10f$finalModel)

# Variable importance from the CV model
# varImp(olr_model_10f)
# plot(varImp(olr_model_10f), top = 10)

# Confusion matrix from CV predictions
cm <- confusionMatrix(olr_model_10f$pred$pred, olr_model_10f$pred$obs)
print(cm)
print.cfM(cm$table, file="cfM-olr-10f.pdf")

library(car)
library(ggplot2)
library(dplyr)

# Get Anova results
anova_results <- Anova(olr_model_10f$finalModel, type = "II")

# Create data frame with variable names and importance
importance_df <- data.frame(
  FullTerm = rownames(anova_results),
  Importance = anova_results$`LR Chisq`,
  P_value = anova_results$`Pr(>Chisq)`
)

# Clean variable names: remove backticks AND polynomial terms
importance_df$BaseVariable <- gsub("`", "", importance_df$FullTerm)  # Remove backticks first
importance_df$BaseVariable <- gsub("\\.L|\\.Q|\\.C|\\^[0-9]+", "", importance_df$BaseVariable)  # Then remove polynomial terms

# Aggregate by base variable (sum the chi-square statistics)
combined_importance <- importance_df %>%
  group_by(BaseVariable) %>%
  summarise(
    TotalImportance = sum(Importance),
    MinPValue = min(P_value),
    NumTerms = n(),  # Number of polynomial terms combined
    .groups = 'drop'
  ) %>%
  arrange(desc(TotalImportance))

# Plot combined importance
top_n <- 10
ggplot(combined_importance[1:top_n, ], 
       aes(x = reorder(BaseVariable, TotalImportance), y = TotalImportance)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "Variable", y = "Combined LR Chi-square", 
       title = paste("Top", top_n, "Variables (Combined Polynomial Terms)")) +
  theme_minimal()

# Show the cleaned results
print(head(combined_importance, 15))
```


```{r}
library(car)
library(ggplot2)
library(dplyr)

# Get Anova results
anova_results <- Anova(olr_model_10f$finalModel, type = "II")

# Create data frame with variable names and importance
importance_df <- data.frame(
  FullTerm = rownames(anova_results),
  Importance = anova_results$`LR Chisq`,
  Df = anova_results$Df,
  P_value = anova_results$`Pr(>Chisq)`
)

# Clean variable names: remove backticks AND polynomial terms
importance_df$BaseVariable <- gsub("`", "", importance_df$FullTerm)
importance_df$BaseVariable <- gsub("\\.L|\\.Q|\\.C|\\^[0-9]+", "", importance_df$BaseVariable)

# Aggregate by base variable - sum chi-square AND degrees of freedom
combined_importance <- importance_df %>%
  group_by(BaseVariable) %>%
  summarise(
    TotalLRChisq = sum(Importance),
    TotalDf = sum(Df),
    MinPValue = min(P_value),
    NumTerms = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    # Calculate proper p-value using combined chi-square and df
    CombinedPValue = pchisq(TotalLRChisq, df = TotalDf, lower.tail = FALSE)
  ) %>%
  arrange(desc(TotalLRChisq))

# Show results
print(head(combined_importance, 15))
```
```{r}
# Using the combined_importance from Option 1
combined_importance$Significance <- ifelse(combined_importance$CombinedPValue < 0.001, "p < 0.001",
                                    ifelse(combined_importance$CombinedPValue < 0.01, "p < 0.01", 
                                    ifelse(combined_importance$CombinedPValue < 0.05, "p < 0.05", "p ≥ 0.05")))

top_n <- 10
ggplot(combined_importance[1:top_n, ], 
       aes(x = reorder(BaseVariable, TotalLRChisq), y = TotalLRChisq, fill = Significance)) +
  geom_col() +
  coord_flip() +
  labs(x = "Variable", y = "Combined LR Chi-square", 
       title = paste("Top", top_n, "Most Important Variables"),
       fill = "Significance Level") +
  scale_fill_manual(values = c("p < 0.001" = "darkgreen", "p < 0.01" = "green", 
                               "p < 0.05" = "orange", "p ≥ 0.05" = "red")) +
  theme_minimal()
```

Calculate Spearman correlation (better for ordinal data)
but Pearson correlation is actually almost the same

```{r}
# Get CV predictions
cv_predictions <- olr_model_10f$pred

spearman_cor <- cor(as.numeric(cv_predictions$pred), as.numeric(cv_predictions$obs), 
                    method = "spearman")
print(paste("Spearman correlation:", round(spearman_cor, 3)))

pearson_cor <- cor(as.numeric(cv_predictions$pred), as.numeric(cv_predictions$obs), 
                    method = "pearson")
print(paste("Pearson  correlation:", round(pearson_cor, 3)))


# Calculate for each fold
library(dplyr)
fold_correlations <- cv_predictions %>%
  group_by(Resample) %>%
  summarise(
    Spearman = cor(as.numeric(pred), as.numeric(obs), method = "spearman"),
    Pearson = cor(as.numeric(pred), as.numeric(obs), method = "pearson"),
    .groups = 'drop'
  )

print("Correlations by fold:")
print(fold_correlations)
print(paste("Mean Spearman correlation:", round(mean(fold_correlations$Spearman), 3)))
print(paste("Mean Pearson correlation:", round(mean(fold_correlations$Pearson), 3)))
```

Try a model with just the top 5 most important variables
(as we will see it has worse performance)

```{r}
train_control<- trainControl(method="cv", number=10, savePredictions = TRUE)
simple_formula <- krize_rep ~ vek + obor_dotovany + kar_praxe + kar_citace + kar_titul
simple_olr_model <- train(simple_formula, data = my, trControl = train_control, method = "polr",
                          na.action = na.omit, tuneGrid = data.frame(method = "logistic"))
# Get CV predictions
cv_predictions <- simple_olr_model$pred

# Calculate Spearman correlation (better for ordinal data)
spearman_cor <- cor(as.numeric(cv_predictions$pred), as.numeric(cv_predictions$obs), 
                    method = "spearman")
print(paste("Spearman correlation:", round(spearman_cor, 3)))
pearson_cor <- cor(as.numeric(cv_predictions$pred), as.numeric(cv_predictions$obs), 
                    method = "pearson")
print(paste("Pearson  correlation:", round(pearson_cor, 3)))


# Calculate for each fold
library(dplyr)
fold_correlations <- cv_predictions %>%
  group_by(Resample) %>%
  summarise(
    Spearman = cor(as.numeric(pred), as.numeric(obs), method = "spearman"),
    Pearson = cor(as.numeric(pred), as.numeric(obs), method = "pearson"),
    .groups = 'drop'
  )

print("Correlations by fold:")
print(fold_correlations)
print(paste("Mean Spearman correlation:", round(mean(fold_correlations$Spearman), 3)))
print(paste("Mean Pearson correlation:", round(mean(fold_correlations$Pearson), 3)))

# Confusion matrix from CV predictions
cm <- confusionMatrix(simple_olr_model$pred$pred, simple_olr_model$pred$obs)
print(cm)
print.cfM(cm$table, file="cfM-simple-olr-10f.pdf")


```

Multinomial Regression using glmnet package 

```{r}
# Create model matrix for glmnet
X <- model.matrix(terms(model_frame_used), model_frame_used)[, -1]
# Fit glmnet 
library(glmnet)
cv_glmnet <- cv.glmnet(X, actual_y, family = "multinomial", alpha = 1)
glmnet_model <- glmnet(X, actual_y, family = "multinomial", lambda = cv_glmnet$lambda.min)
# Get model coefficients
coef(glmnet_model)

```



```{r}
  # 1. CLASS PREDICTIONS (most common for confusion matrix)
  class_pred <- predict(glmnet_model, newx = X, s = cv_glmnet$lambda.min, type = "class")
  class_pred_factor <- as.factor(class_pred[, 1])
  # Result: matrix with 1 column containing factor levels
  # 2. PROBABILITY PREDICTIONS
  prob_pred <- predict(glmnet_model, newx = X, s = cv_glmnet$lambda.min, type = "response")
  prob_matrix <- prob_pred[, , 1]  # Extract the probability matrix
  # Result: array with probabilities for each class
  # Get class names
  class_names <- dimnames(prob_matrix)[[2]]
  # compare predictions with actual values in data
  library(caret)
  class_pred_factor <- factor(class_pred_factor, levels = levels(actual_y))
  cm <- confusionMatrix(data = class_pred_factor, reference = as.factor(actual_y))
  print.cfM(cm$table, file="cfM-glmnet.pdf")
```

```{r}
dep.var.levels <- names(table(my[,dependent.var]))
strong.vars <- c()
for (level in dep.var.levels){
  vars <- rownames(glmnet_model$beta[[level]])[which(abs(glmnet_model$beta[[level]])>0.3)]
  strong.vars <- unique(c(strong.vars, vars))
}
print(strong.vars)
```

```{r}
library(dplyr)
library(randomForest)

source("functions.R")
# Apply the importance functions
var_importance_all <- get_variable_importance(multinom_model=model, glmnet_model=glmnet_model, 
                                              X=X, y=actual_y, original_vars=good_vars)
print("Comprehensive Variable Importance:")
print(var_importance_all[, c("Variable", "Composite_Importance", "Multinom_Importance", "Glmnet_Importance", "RF_Importance")])

# Group by original variables
grouped_importance <- group_variable_importance(
  data.frame(Variable = var_importance_all$Variable, 
            Importance = var_importance_all$Composite_Importance),
  all_predictors
)

print("\nGrouped by Original Variables:")
print(grouped_importance)

```

Cluster analysis

```{r}
library(mclust)

# Remove variables with excessive missing values (e.g., >20% missing)
threshold <- 0.1
my.vars.clean <- my.vars.for_cor[, sapply(my.vars.for_cor, function(x) mean(is.na(x))) < threshold]

# Then remove remaining incomplete cases
complete_data <- na.omit(my.vars.clean)
# nrow(complete_data)

mod1 <- Mclust(complete_data)
summary(mod1)
plot(mod1,  what = "BIC")
```

```{r}
# 1. Basic cluster summary
summary(mod1)

# 2. Cluster means comparison
cluster_means <- mod1$parameters$mean
print(cluster_means)

# 3. Cluster sizes
cluster_sizes <- table(mod1$classification)
print(cluster_sizes)

# 4. Create detailed comparison
library(dplyr)
# Add cluster assignments to your original data
clustered_data <- cbind(complete_data, cluster = mod1$classification)

# Calculate descriptive statistics by cluster
cluster_summary <- clustered_data %>%
  group_by(cluster) %>%
  summarise_all(list(mean = ~mean(., na.rm=TRUE), 
                    sd = ~sd(., na.rm=TRUE)))

# 5. Visualize differences
plot(mod1, what = "classification")
```

Looking at the cluster means, here's what characterizes each group:

## Cluster Descriptions:

### **Cluster 1 (n=284): "High-Profile Academic Cluster"**
This cluster is characterized by **higher values** in:
- **obor_medializ** (4.20 vs 1.84) - Much more media attention
- **obor_publikac** (4.14 vs 3.39) - More publication-oriented
- **obor_dotovany** (3.05 vs 2.24) - Better funded fields
- **obor_konkuren** (4.11 vs 3.44) - More competitive fields
- **obor_priliv** (3.01 vs 2.14) - More attractive/popular fields
- **obor_rychly** (3.44 vs 2.88) - Faster-moving fields
- **obor_komplexn** (4.26 vs 3.87) - More complex fields
- **kar_praxe** (3.03 vs 2.90) - Slightly more practical experience
- **kar_citace** (2.44 vs 2.28) - Slightly more citations

### **Cluster 2 (n=446): "Lower-Profile Academic Cluster"** 
This cluster is characterized by **higher values** in:
- **krize_rep** (2.30 vs 1.95) - More reputation crises/challenges
- **obor_mensi** (2.81 vs 2.33) - Smaller fields
- **pomer_zakladni** (8.05 vs 7.90) - Slightly higher basic proportion
- **kar_nejlepsi** (1.74 vs 1.68) - Slightly better career outcomes
- **cinnost_exper** (1.75 vs 1.63) - More experimental activities

## Key Insights:

1. **Cluster 1** represents academics in **high-visibility, well-funded, competitive fields** with more media attention and publication focus.

2. **Cluster 2** represents academics in **smaller, less visible fields** that may face more reputational challenges but have slightly better individual outcomes.

3. **Gender** is nearly identical across clusters (1.69 vs 1.67), suggesting this clustering isn't gender-driven.

The biggest discriminating factors are **media attention (obor_medializ)**, **funding (obor_dotovany)**, and **field attractiveness (obor_priliv)** - essentially separating "high-profile" from "low-profile" academic fields.

**************************************************************************************************************

Latent Class Analysis (more suitable than cluster analysis for nominal data in spec_hlavni)

```{r}
library(poLCA)

# if we want to consider also this variable with 10 values
my.vars.bn <- cbind(my.vars.for_cor, my_clean$spec_hlavni)
colnames(my.vars.bn) <- c(colnames(my.vars.bn)[-length(colnames(my.vars.bn))],"spec_hlavni")
# Convert all columns to factors at once
my.vars.bn.discrete <- as.data.frame(lapply(my.vars.bn, as.factor))

f <- create_polca_formula(my.vars.bn)
min_bic <- Inf
bic.values <- rep(0,8-2+1)
for(i in 2:8){
  lc <- poLCA(f, data = my.vars.bn.discrete, nclass=i, maxiter=3000, 
              tol=1e-5, na.rm=FALSE,  
              nrep=10, verbose=FALSE, calc.se=TRUE)
  bic.values[i-1] <- lc$bic 
  if(lc$bic < min_bic){
    min_bic <- lc$bic
    LCA_best_model<-lc
  }
}    	
LCA_best_model
# conditional probabilities
LCA_best_model$probs$B1
# ...
LCA_best_model$probs$C16
# prior class probabilities
LCA_best_model$P

pdf(file="LCA-BIC-values.pdf", width=5,height=4)
  plot(x=2:8,y=-bic.values, xlab="number of latent classes", ylab="BIC value",type="l", lwd=2)
  grid()
dev.off()
```

Find variables most related to the latent class

```{r}
library(DescTools)
lng <- length(LCA_best_model$probs)
rr <- rep(0,lng)
names(rr) <- names(LCA_best_model$probs)
for (i in 1:lng){
  cpt <- LCA_best_model$probs[i][[1]]
  pt  <- cpt * matrix(rep(LCA_best_model$P,ncol(cpt)),nrow=3,ncol=ncol(cpt))
  rr[i] <- CramerV(pt)
}
sort(rr,decreasing = TRUE)

pc <- matrix(0,ncol=length(LCA_best_model$P),nrow=lng)
rownames(pc) <- names(LCA_best_model$probs)
colnames(pc) <- rownames(LCA_best_model$probs[1][[1]])
sc <- pc
for (i in 1:lng){
  cpt <- LCA_best_model$probs[i][[1]]
  pc[i,] <- apply(cpt,1,max)
  sc[i,] <- apply(cpt,1,which.max)
}
pco <- matrix(0,ncol=length(LCA_best_model$P),nrow=lng)
colnames(pco) <- rownames(LCA_best_model$probs[1][[1]])
pco <- pc
sco <- sc
varo <- pc
for (c in 1:ncol(pc)){
  o <- order(pc[,c],decreasing = TRUE)
  pco[,c] <- pc[,c][o]
  sco[,c] <- sc[,c][o]
  varo[,c] <- rownames(pc)[o]
}
rownames(pco) <- NULL
rownames(sco) <- NULL
rownames(varo) <- NULL
```


```{r}
# Calculate actual values from your poLCA results for bubble plot
# LCA_best_model$
# Assuming your poLCA model is stored as 'mod1' or similar

# 1. EXPERIMENTAL INTENSITY (cinnost_exper) - weighted average for each class
# LCA_best_model$probs$cinnost_exper has levels 1,2,3,4
# experimental_intensity <- c(
#   # Class 1: 0.6680*1 + 0.1808*2 + 0.1210*3 + 0.0302*4
#   0.6680*1 + 0.1808*2 + 0.1210*3 + 0.0302*4,
#   
#   # Class 2: 0.5374*1 + 0.2241*2 + 0.1109*3 + 0.1276*4  
#   0.5374*1 + 0.2241*2 + 0.1109*3 + 0.1276*4,
#   
#   # Class 3: 0.4448*1 + 0.1894*2 + 0.1275*3 + 0.2383*4
#   0.4448*1 + 0.1894*2 + 0.1275*3 + 0.2383*4
# )
experimental_intensity <- LCA_best_model$probs$cinnost_exper %*% 1:4

print("Experimental Intensity by Class:")
print(paste("Class 1:", round(experimental_intensity[1], 2)))
print(paste("Class 2:", round(experimental_intensity[2], 2)))
print(paste("Class 3:", round(experimental_intensity[3], 2)))

# 2. CRISIS PERCEPTION (krize_rep) - weighted average for each class
# From your output: krize_rep has levels 1,2,3,4
# crisis_perception <- c(
#   # Class 1: 0.5111*1 + 0.2997*2 + 0.1181*3 + 0.0711*4
#   0.5111*1 + 0.2997*2 + 0.1181*3 + 0.0711*4,
#   
#   # Class 2: 0.2932*1 + 0.2530*2 + 0.3686*3 + 0.0851*4
#   0.2932*1 + 0.2530*2 + 0.3686*3 + 0.0851*4,
#   
#   # Class 3: 0.2169*1 + 0.2925*2 + 0.2675*3 + 0.2231*4
#   0.2169*1 + 0.2925*2 + 0.2675*3 + 0.2231*4
# )
crisis_perception <- LCA_best_model$probs$krize_rep %*% 1:4

print("\nCrisis Perception by Class:")
print(paste("Class 1:", round(crisis_perception[1], 2)))
print(paste("Class 2:", round(crisis_perception[2], 2)))
print(paste("Class 3:", round(crisis_perception[3], 2)))

# 3. CLASS SIZES (from estimated class population shares)
# class_sizes <- c(0.2669, 0.473, 0.2601)  # From your output
class_sizes <- LCA_best_model$P
total_n <- nrow(my.vars.bn.discrete) # Total observations
actual_class_sizes <- class_sizes * total_n

print("\nClass Sizes:")
print(paste("Class 1:", round(actual_class_sizes[1])))
print(paste("Class 2:", round(actual_class_sizes[2])))
print(paste("Class 3:", round(actual_class_sizes[3])))

# 4. FIELD DISTRIBUTIONS by class (from spec_hlavni)
# Create field groupings for cleaner visualization
biology_prop <- c(0.5336, 0.3678, 0.2417)  # Biology proportion in each class
biology_prop <- LCA_best_model$probs$spec_hlavni[,"Biology"]

physics_earth_prop <- c(
  0.0082 + 0.0393 + 0.0952,  # Class 1: Astronomy + Earth + Physics
  0.0050 + 0.1413 + 0.1230,  # Class 2: Astronomy + Earth + Physics  
  0.0175 + 0.1274 + 0.2042   # Class 3: Astronomy + Earth + Physics
)
chem_materials_prop <- LCA_best_model$probs$spec_hlavni[,"Astronomy and Planetary Science"]
+LCA_best_model$probs$spec_hlavni[,"Earth and Environmental Science"]+LCA_best_model$probs$spec_hlavni[,"Physics"]   
# chem_materials_prop <- c(0.1819 + 0.0293, 0.1254 + 0.0979, 0.1238 + 0.0722)

# Create comprehensive data frame for bubble plot
bubble_data_calculated <- data.frame(
  Class = c("Class 1: Elite", "Class 2: Early Career", "Class 3: Senior Exp"),
  Crisis_Perception = crisis_perception,
  Experimental_Intensity = experimental_intensity,
  Class_Size = actual_class_sizes,
  Biology_Prop = biology_prop,
  Physics_Earth_Prop = physics_earth_prop,
  Chem_Materials_Prop = chem_materials_prop
)

print("\nFinal Data for Bubble Plot:")
print(bubble_data_calculated)

# 5. Create the actual bubble plot with calculated values
library(ggplot2)

# Simple version with one bubble per class
p1 <- ggplot(bubble_data_calculated, aes(x = Experimental_Intensity, 
                                        y = Crisis_Perception,
                                        size = Class_Size)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  geom_text(aes(label = gsub("Class [0-9]: ", "", Class)), 
            vjust = -1.5, size = 3.5, show.legend = FALSE) +
  
  scale_size_continuous(name = "Class Size", range = c(5, 20)) +
  
  labs(title = "Reproducibility Crisis Perception vs Experimental Intensity",
       subtitle = "Each bubble represents one latent class of academics",
       x = "Experimental Intensity (1=low, 4=high)",
       y = "Reproducibility Crisis Concern (1=no crisis, 4=severe crisis)",
       caption = "Bubble size = number of academics in class") +
  
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 11)) +
  
  # Add correlation line
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed", size = 0.8) +
  
  # Expand axes slightly for better label visibility  
  expand_limits(x = c(1, 2.5), y = c(1.5, 2.8))

pdf(file="LCA Reproducibility Crisis Perception vs Experimental Intensity.pdf")
  print(p1)
dev.off()

# 6. Enhanced version showing field composition
# Reshape data to show field breakdowns
library(tidyr)
library(dplyr)

# Check column names first
print("Column names in bubble_data_calculated:")
print(names(bubble_data_calculated))

# Alternative approach - create field data manually to avoid select issues
field_data <- data.frame(
  Class = rep(c("Class 1: Elite", "Class 2: Early Career", "Class 3: Senior Exp"), each = 3),
  Crisis_Perception = rep(crisis_perception, each = 3),
  Experimental_Intensity = rep(experimental_intensity, each = 3),
  Class_Size = rep(actual_class_sizes, each = 3),
  Field = rep(c("Biology", "Physics/Earth", "Chemistry/Materials"), 3),
  Proportion = c(
    biology_prop[1], physics_earth_prop[1], chem_materials_prop[1],  # Class 1
    biology_prop[2], physics_earth_prop[2], chem_materials_prop[2],  # Class 2  
    biology_prop[3], physics_earth_prop[3], chem_materials_prop[3]   # Class 3
  )
) %>%
  mutate(Field_Size = Class_Size * Proportion) %>%
  filter(Proportion > 0.05)  # Only show substantial field representations

# BETTER OPTION 1: Side-by-side stacked bar charts
crisis_field_data <- data.frame(
  Class = rep(c("Class 1:\nElite Math", "Class 2:\nEarly Career", "Class 3:\nSenior Exp"), each = 4),
  Crisis_Level = rep(c("No Crisis", "Low Crisis", "Moderate Crisis", "High Crisis"), 3),
  Probability = c(
    # Class 1 probabilities (from krize_rep output)
    0.51, 0.30, 0.12, 0.07,
    # Class 2 probabilities  
    0.29, 0.25, 0.37, 0.09,
    # Class 3 probabilities
    0.22, 0.29, 0.27, 0.22
  ),
  # Add experimental intensity for positioning
  Experimental_Intensity = rep(experimental_intensity, each = 4)
) %>%
  mutate(Crisis_Level = factor(Crisis_Level, 
                              levels = c("No Crisis", "Low Crisis", "Moderate Crisis", "High Crisis")))

p2_stacked <- ggplot(crisis_field_data, aes(x = reorder(Class, Experimental_Intensity), 
                                           y = Probability, 
                                           fill = Crisis_Level)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(name = "Crisis Perception",
                    values = c("No Crisis" = "#2E8B57",
                              "Low Crisis" = "#90EE90", 
                              "Moderate Crisis" = "#FFD700",
                              "High Crisis" = "#DC143C")) +
  labs(title = "Reproducibility Crisis Perceptions by Academic Class",
       subtitle = "Classes ordered by experimental intensity (left=least, right=most experimental)",
       x = "Academic Class",
       y = "Proportion",
       caption = "Clear trend: More experimental → More crisis concern") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10),
        legend.position = "bottom")

pdf(file = "LCA Reproducibility Crisis Perceptions by Academic Class.pdf")
  print(p2_stacked)
dev.off()

# BETTER OPTION 2: Faceted field composition plot  
field_summary <- data.frame(
  Class = rep(c("Elite Math", "Early Career", "Senior Exp"), 3),
  Field = rep(c("Biology", "Physics/Earth", "Chemistry/Materials"), each = 3),
  Proportion = c(biology_prop, physics_earth_prop, chem_materials_prop),
  Crisis_Perception = rep(crisis_perception, 3),
  Experimental_Intensity = rep(experimental_intensity, 3)
) %>%
  filter(Proportion > 0.1)  # Only show fields with >10% representation

p3_faceted <- ggplot(field_summary, aes(x = Experimental_Intensity, 
                                       y = Crisis_Perception,
                                       size = Proportion)) +
  geom_point(alpha = 0.8, color = "steelblue") +
  
  # Use repel to avoid overlapping and keep labels inside plot
  ggrepel::geom_text_repel(aes(label = Class), 
                          size = 2.8, 
                          show.legend = FALSE,
                          max.overlaps = 20,
                          box.padding = 0.3,
                          point.padding = 0.2) +
  
  facet_wrap(~ Field, scales = "free") +
  
  scale_size_continuous(name = "Field Proportion\nwithin Class", 
                       range = c(4, 12), 
                       labels = scales::percent) +
  
  labs(title = "Crisis Perception vs Experimental Intensity by Field",
       subtitle = "Each panel shows one field; bubble size = proportion of class in that field",
       x = "Experimental Intensity",
       y = "Crisis Perception") +
  
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        strip.text = element_text(size = 11, face = "bold"),
        panel.spacing = unit(1, "lines")) +
  
  # Expand plot margins to accommodate labels
  coord_cartesian(clip = "off")

pdf(file="LCA Crisis Perception vs Experimental Intensity by Field.pdf")
  print(p3_faceted)
dev.off()

# BEST OPTION 3: Clean correlation plot with annotations
correlation_data <- data.frame(
  Class = c("Elite Math", "Early Career", "Senior Exp"),
  Crisis_Perception = crisis_perception,
  Experimental_Intensity = experimental_intensity,
  Class_Size = actual_class_sizes,
  Top_Field = c("Biology (53%)", "Biology (37%)", "Physics (20%)")
)

p4_clean <- ggplot(correlation_data, aes(x = Experimental_Intensity, 
                                        y = Crisis_Perception,
                                        size = Class_Size)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  
  # Add class labels and top field info
  geom_text(aes(label = paste0(Class, "\n", Top_Field)), 
            vjust = -1.8, hjust = 0.5, size = 3.5, 
            lineheight = 0.8, show.legend = FALSE) +
  
  # Correlation line (inherit.aes = FALSE to avoid size aesthetic warning)
  geom_smooth(aes(x = Experimental_Intensity, y = Crisis_Perception), 
              method = "lm", se = TRUE, color = "red", 
              linetype = "dashed", alpha = 0.3, inherit.aes = FALSE) +
  
  scale_size_continuous(name = "Class Size", range = c(8, 20)) +
  
  labs(title = "The Reproducibility Paradox",
       subtitle = "More experimental academics are more concerned about reproducibility crisis",
       x = "Experimental Intensity →",
       y = "Crisis Concern →",
       caption = "Bubble size = number of academics; Top field shown for each class\nCorrelation confirms: experimental experience drives crisis awareness") +
  
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12),
        plot.caption = element_text(size = 10, color = "gray60"),
        axis.title = element_text(size = 11)) +
  
  # Expand limits for better label visibility
  expand_limits(x = c(1.3, 2.4), y = c(1.6, 2.8))

pdf(file="LCA Experimental academics are more concerned about reproducibility crisis.pdf")
  print(p4_clean)
dev.off()
```

Claude's comment to the output
**krize_rep** measures academics' **perceptions of the reproducibility crisis**, not personal reputation issues. 

## **Reproducibility Crisis Perceptions by Class:**

### **Class 1 (Elite Math Academics): LEAST concerned about reproducibility crisis**
- **51% see NO/minimal crisis** (level 1) - most optimistic
- Only 19% see moderate-severe crisis (levels 3-4)

### **Class 2 (Early Career Mainstream): MODERATE concern**  
- **29% see no crisis** - more aware than elites
- **45% see moderate-severe crisis** (levels 3-4)

### **Class 3 (Senior Experimentalists): MOST concerned about reproducibility crisis**
- **Only 22% dismiss the crisis** - most pessimistic group
- **49% see moderate-severe crisis** (levels 3-4)
- **22% see severe crisis** (level 4) - twice other classes!

## **Field-Specific Reproducibility Concerns:**

### **HIGH-CONCERN Fields:**
- **Physics** (Class 3: 20%) - complex experiments, instrumentation issues
- **Earth/Environmental Science** (Classes 2&3: ~13-14%) - modeling challenges, data complexity

### **LOW-CONCERN Fields:**
- **Biology** (Class 1: 53%) - despite known reproducibility issues, these academics are less worried
- **Mathematics** - theoretical work less affected by reproducibility problems

## **Key Insights:**

1. **"Experience Effect"**: Senior experimentalists (Class 3) have **seen the problems firsthand** - they know how hard reproducibility is

2. **"Mathematical Blind Spot"**: Elite mathematical academics (Class 1) may be **insulated** from reproducibility issues in their theoretical work

3. **"Experimental Reality"**: Classes doing more experimental work (especially Class 3) are **acutely aware** of reproducibility challenges

4. **"Field Sensitivity"**: Physics and Earth Sciences practitioners are more concerned - fields with complex experiments and models

## **Practical Implications:**

- **Experimental scientists** are more realistic about reproducibility challenges
- **Theoretical/mathematical** researchers may underestimate the scope of the crisis
- **Senior academics** with experimental experience are most **alarmed** - they've witnessed the problems evolve
- **Early career** researchers have moderate awareness but less direct experience

This suggests reproducibility crisis awareness follows **direct exposure** - those doing the experiments know the problems best, while theoretical researchers may not fully grasp the extent of the issue.

*****************************************************************************************************************************
Claude's comment to the output without spec_hlavni:
Looking at this poLCA output, here's my interpretation of the 3-class model:

## Class Characteristics

**Class 1 (25.6% of sample): "Senior Established Researchers"**
- **Demographics**: Older (high probability in age categories 3-6), mostly male (67.3%), more experienced
- **Career**: High citations (balanced across categories), senior positions (kar_praxe: 42.2% category 3, 38.1% category 4)
- **Research style**: More mathematical/modeling work, less experimental
- **Field characteristics**: Work in highly funded, competitive, high-pressure fields

**Class 2 (47.6% of sample): "Junior/Mid-career Researchers"** 
- **Demographics**: Younger (93.9% in age category 3), mixed gender (58.2% male)
- **Career**: Lower citations (50.2% in lowest category), earlier career stages
- **Research style**: Less mathematical, more varied activities
- **Field characteristics**: Work in less competitive, more accessible fields

**Class 3 (26.8% of sample): "Elite Senior Researchers"**
- **Demographics**: Oldest group, predominantly male (84.8%)
- **Career**: Highest citations and seniority (88.4% in top kar_praxe category)
- **Research style**: Mixed mathematical/experimental approaches
- **Field characteristics**: Highly competitive and well-funded fields

## Key Distinguishing Variables

**Most important variables for class differentiation:**

1. **Age (vek)**: Clear age stratification across classes
2. **Career experience (kar_praxe)**: Strong predictor of class membership  
3. **Citations (kar_citace)**: Classes 1 & 3 have higher citation patterns
4. **Field funding (obor_dotovany)**: Class 1 works in highly funded fields
5. **Mathematical activity (cinnost_matem)**: Higher in senior classes

## Relation to Crisis of Representation (krize_rep)

**This is the crucial finding for your analysis:**

- **Class 1**: 50.8% report no crisis, 30.2% mild crisis → **Lower crisis perception**
- **Class 2**: More evenly distributed (29.8% no crisis, 36.7% moderate crisis) → **Higher crisis awareness** 
- **Class 3**: Most distributed across all levels (22.5% no crisis, 22.0% severe crisis) → **Mixed/polarized views**

## Key Interpretations

**The crisis of representation varies significantly by researcher profile:**

1. **Senior established researchers** (Class 1) are less likely to perceive a representation crisis
2. **Junior/mid-career researchers** (Class 2) are most aware of representation issues
3. **Elite senior researchers** (Class 3) show the most polarized views - some see no crisis while others see it as severe

**This suggests that career stage and institutional position influence perception of representation crises.** The middle-career group may be most sensitive to these issues, possibly because they're actively navigating career advancement and facing these pressures directly.

The model fit statistics suggest this is a reasonable solution, though the large sample size makes goodness-of-fit tests less reliable. The clear differentiation in krize_rep across classes indicates this latent class structure captures meaningful variation in your key outcome variable.
************************************************************************************************************************************

PCA for "obor_"

```{r}

PCA.vars <- c("obor_dotovany",
"obor_konkuren",
"obor_tlak",
"obor_rychly",
"obor_priliv",
"obor_publikac",
"obor_medializ",
"obor_verejnos",
"obor_komplexn",
"obor_mensi",
"obor_ekzaktnost"
)

library(missMDA)
library(FactoMineR)

# This is often the most robust approach
# Step 1: Estimate optimal number of components
nb_components <- estim_ncpPCA(my.vars.for_cor[,PCA.vars], 
                              method.cv = "Kfold", verbose = FALSE)

# Step 2: Impute missing values using PCA
imputed_result <- imputePCA(my.vars.for_cor[,PCA.vars], 
                           ncp = nb_components$ncp)

# Step 3: Run final PCA
pca_result <- PCA(imputed_result$completeObs, 
                 scale.unit = TRUE, 
                 graph = FALSE)

# Access results
# pca_result$ind$coord    # Individual coordinates (scores)
pca_result$var$coord    # Variable coordinates (loadings)
pca_result$eig          # Eigenvalues and explained variance

# pca_result <- prcomp(my.vars.for_cor[,PCA.vars], scale. = TRUE)


```

PCA plots

```{r}
# Assuming the PCA result is stored in 'pca_result'

library(ggplot2)
library(ggrepel)

# Convert to data frame
loadings_df <- data.frame(
  Variable = rownames(pca_result$var$coord),
  PC1 = pca_result$var$coord[,1],
  PC2 = pca_result$var$coord[,2]
)

# Plot loadings for first two components
pdf(file="fields-PC1-PC2.pdf")
  ggplot(loadings_df, aes(x = PC1, y = PC2)) +
    geom_point(color = "blue", size = 3) +
    geom_text_repel(aes(label = Variable), 
                    size = 3,
                    box.padding = 0.5,      # Space around labels
                    point.padding = 0.3,    # Space from points
                    max.overlaps = Inf) +   # Allow all labels
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
    labs(x = paste0("PC1 (", round(pca_result$eig[1,2], 1), "%)"),
         y = paste0("PC2 (", round(pca_result$eig[2,2], 1), "%)"),
         title = "PCA - Variables Loadings") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
dev.off()
```
```{r}

library(ggplot2)
library(gridExtra)

# Extract eigenvalues and variance
eigenvals <- pca_result$eig[,1]
var_explained <- pca_result$eig[,2]
cum_var <- pca_result$eig[,3]

pdf(file="fields-scree-cum-plots.pdf", width = 14, height = 7)
  # Scree plot
  scree_plot <- ggplot(data.frame(PC = 1:length(eigenvals), Eigenvalue = eigenvals), 
                       aes(x = PC, y = Eigenvalue)) +
    geom_line(color = "blue", size = 1) +
    geom_point(color = "blue", size = 3) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
    labs(title = "Scree Plot", x = "Principal Component", y = "Eigenvalue") +
    theme_minimal()
  
  # Cumulative variance plot  
  cum_plot <- ggplot(data.frame(PC = 1:length(cum_var), CumVar = cum_var), 
                     aes(x = PC, y = CumVar)) +
    geom_line(color = "green", size = 1) +
    geom_point(color = "green", size = 3) +
    geom_hline(yintercept = 80, linetype = "dashed", color = "red") +
    labs(title = "Cumulative Variance", x = "Principal Component", y = "Cumulative % Variance") +
    theme_minimal()
  
  grid.arrange(scree_plot, cum_plot, ncol = 2)
dev.off()
```
## Key Interpretations from PCA results:

**1. Kaiser Criterion (Eigenvalue > 1.0):**
- Components 1, 2, and 3 have eigenvalues > 1.0
- **Suggests 3 components** as optimal

**2. Scree Plot Analysis:**
- Large drop from PC1 (3.48) to PC2 (1.09) - major "elbow"
- Gradual decline afterward with a smaller elbow around PC3-PC4
- **Suggests 2-3 components**

**3. Variance Explained:**
- PC1 alone explains 31.6% of variance (dominant component)
- First 3 components explain 50.6% of total variance
- First 7 components needed for 80% variance threshold

## **My Recommendation:**

Based on your data, I'd suggest **3 components** as optimal because:
- Meets Kaiser criterion (eigenvalue > 1.0)
- Captures over 50% of total variance
- Shows clear elbow in scree plot after PC3
- Maintains good interpretability


```{r}
library(factoextra)

# Circle of correlations
fviz_pca_var(pca_result, 
             axes = c(1,2),            # Which PCs to plot
             col.var = "black",
             alpha.var = 0.7,
             repel = TRUE) +
  theme_minimal() +
  ggtitle("PCA - Circle of Correlations")

# For different PC pairs
fviz_pca_var(pca_result, axes = c(1,3))  # PC1 vs PC3
fviz_pca_var(pca_result, axes = c(2,3))  # PC2 vs PC3

```
Practical Implications:

PC1 helps you identify which fields are most "active" overall
PC2 distinguishes between technical fields (mathematics, physics) vs. public fields (journalism, politics)
PC3 reveals the relationship between field size and competitive pressure

***************************************************************************************************************************

PCA for "kar_"

```{r}
PCA.vars <- c("kar_praxe",
"kar_titul",
"kar_citace",
"kar_nejlepsi")

library(missMDA)
library(FactoMineR)

# This is often the most robust approach
# Step 1: Estimate optimal number of components
nb_components <- estim_ncpPCA(my.vars.for_cor[,PCA.vars], 
                              method.cv = "Kfold", verbose = FALSE)

# Step 2: Impute missing values using PCA
imputed_result <- imputePCA(my.vars.for_cor[,PCA.vars], 
                           ncp = nb_components$ncp)

# Step 3: Run final PCA
pca_result <- PCA(imputed_result$completeObs, 
                 scale.unit = TRUE, 
                 graph = FALSE)

# Access results
# pca_result$ind$coord    # Individual coordinates (scores)
pca_result$var$coord    # Variable coordinates (loadings)
pca_result$eig          # Eigenvalues and explained variance

# pca_result <- prcomp(my.vars.for_cor[,PCA.vars], scale. = TRUE)

```
```{r}

library(ggplot2)
library(ggrepel)

# Convert to data frame
loadings_df <- data.frame(
  Variable = rownames(pca_result$var$coord),
  PC1 = pca_result$var$coord[,1],
  PC2 = pca_result$var$coord[,2]
)

# Plot loadings for first two components
pdf(file="career-PC1-PC2.pdf")
  ggplot(loadings_df, aes(x = PC1, y = PC2)) +
    geom_point(color = "blue", size = 3) +
    geom_text_repel(aes(label = Variable), 
                    size = 3,
                    box.padding = 0.5,      # Space around labels
                    point.padding = 0.3,    # Space from points
                    max.overlaps = Inf) +   # Allow all labels
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
    labs(x = paste0("PC1 (", round(pca_result$eig[1,2], 1), "%)"),
         y = paste0("PC2 (", round(pca_result$eig[2,2], 1), "%)"),
         title = "PCA - Variables Loadings") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
dev.off()

```


```{r}
library(ggplot2)
library(gridExtra)

# Extract eigenvalues and variance
eigenvals <- pca_result$eig[,1]
var_explained <- pca_result$eig[,2]
cum_var <- pca_result$eig[,3]

pdf(file="career-scree-cum-plots.pdf", width = 14, height = 7)
  # Scree plot
  scree_plot <- ggplot(data.frame(PC = 1:length(eigenvals), Eigenvalue = eigenvals), 
                       aes(x = PC, y = Eigenvalue)) +
    geom_line(color = "blue", size = 1) +
    geom_point(color = "blue", size = 3) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
    labs(title = "Scree Plot", x = "Principal Component", y = "Eigenvalue") +
    theme_minimal()
  
  # Cumulative variance plot  
  cum_plot <- ggplot(data.frame(PC = 1:length(cum_var), CumVar = cum_var), 
                     aes(x = PC, y = CumVar)) +
    geom_line(color = "green", size = 1) +
    geom_point(color = "green", size = 3) +
    geom_hline(yintercept = 80, linetype = "dashed", color = "red") +
    labs(title = "Cumulative Variance", x = "Principal Component", y = "Cumulative % Variance") +
    theme_minimal()
  
  grid.arrange(scree_plot, cum_plot, ncol = 2)
dev.off()
```
```{r}
library(factoextra)

# Circle of correlations
fviz_pca_var(pca_result, 
             axes = c(1,2),            # Which PCs to plot
             col.var = "black",
             alpha.var = 0.7,
             repel = TRUE) +
  theme_minimal() +
  ggtitle("PCA - Circle of Correlations")

```
Key Insight:

Being "nejlepsi" (publishing in best/elite journal) is fundamentally different from having high traditional career metrics. 
This suggests two distinct career success dimensions in your data.
